Multi-domain, multilingual translation with latent multi-task group dropout

### 2022-01-14 2022/acl/acl_latex.tex
revised introduction

In this work, we study and develop new methods aimed to take into account the similarity between tasks in a more effective way, by \emph{learning the network organization from the data}. To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the representation space and show how training can still be performed end-to-end using a variational surrogate of the log-likelihood loss function. Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, meaning that related tasks use the same subparts of the network. Our results also show that this method is comparable to using adapter layers in a number of empirical comparisons; however, contrarily to adapters, these performance are obtained without any increase of the model size. Our main contributions are summarized below:
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
\item We provide a sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
\item We present algorithms to train this model end-to-end with very little extra parameters;
\item We report, using an extensive experimental setting, significant gains of this method with respect to strong baselines for low-resource multilingual MT;
\item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
\end{itemize}
\fyDone{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}

>>>>

In this work, we study and develop a new method, \emph{multi-task group dropout}, aimed to take into account the similarity between tasks in a more effective way, by \emph{learning the network organization from the data}. To this end, we introduce a set of latent variables in the model, to account for the unseen association between tasks and regions of the representation space and show how training can still be performed end-to-end using a variational surrogate of the log-likelihood loss function. Our experiments with multilingual and multidomain machine translation confirm that this method can automatically detect similarities in the data, meaning that related tasks use the same subparts of the network. Our results also show that this method is comparable to using adapter layers in a number of empirical comparisons; however, contrarily to adapters, these performance are obtained without any increase of the model size. Our contributions are primarily methodological and can be summarized as follows:
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
\item We introduce a novel, sound mathematical formulation to the problem of jointly learning task-dependent sub-networks and the parameters of the underlying models using variational probabilistic modeling techniques;
\item We present algorithms to train this model end-to-end with very little extra parameters;
\item We report, using an extensive set of experiments, significant gains with respect to strong baselines for low-resource multilingual MT;
\item We study how this method can actually exploit the similarities between tasks to learn interpretable sub-networks.
\end{enumerate}
\fyDone{Can you do distillation ? That is only keep the parameter for one domain and still get good results (better than DA ?)}
